---
title: "R Notebook"
---

We're interested how much an input set of GO terms spans the GO ontology.

```{r}
library(data.table)
library(ontologyIndex)
library(ontologyPlot)
library(purrr)
library(ggplot2)
library(ggpubr)
library(GO.db)

RESULTS_PATH = '../../data_scAgeCom/analysis/a4_data_results_all_cases.rds'
results = readRDS(RESULTS_PATH)
facs = results$facs_mixed

go = get_ontology(file='../../data/go.obo')
go
```

Check subsets of GO

```{r}
BIOLOGICAL_PROCESS_ROOT = 'GO:0008150'
MOLECULAR_FUNCTION_ROOT = 'GO:0003674'
CELLULAR_COMPONENT_ROOT = 'GO:0005575'

BIOLOGICAL_PROCESS_GOs = get_descendants(go, BIOLOGICAL_PROCESS_ROOT)
message("Number of BP GO terms: ", length(BIOLOGICAL_PROCESS_GOs))
```

Check how much GOs we span

```{r}
# Get union gos across all the tissues
cci_gos = purrr::reduce(
    map(
        names(facs),
        function(.x) {
            unique(facs[[.x]]@ORA$GO[, Value])
        }
    ),
    union
)

message("Number of unique GO terms in CCI: ", length(cci_gos))
message("Number of unique GO BP terms in CCI: ", sum(cci_gos %in% BIOLOGICAL_PROCESS_GOs))

num_unique_bp_gos_per_tiss = map_int(
    names(facs),
    function(.x) {
        sum(
            unique(facs[[.x]]@ORA$GO[, Value])
            %in% BIOLOGICAL_PROCESS_GOs
        )
    }
)

# Num unique bp go in tissue
g = (
    ggplot(data.frame(num_unique_bp_gos_per_tiss), aes(num_unique_bp_gos_per_tiss)) 
    + geom_histogram(binwidth=100)
)
g

# OR_DIFF distribution across all tissues
map(
    names(facs),
    function(.x) {
        
        or_diff = facs[[.x]]@ORA$GO[Value %in% BIOLOGICAL_PROCESS_GOs, .(OR_DIFF)]
        pval_diff = facs[[.x]]@ORA$GO[Value %in% BIOLOGICAL_PROCESS_GOs, .(pval_DIFF)]
        
        or_up = facs[[.x]]@ORA$GO[Value %in% BIOLOGICAL_PROCESS_GOs, .(OR_UP)]
        pval_up = facs[[.x]]@ORA$GO[Value %in% BIOLOGICAL_PROCESS_GOs, .(pval_UP)]
        
        or_down = facs[[.x]]@ORA$GO[Value %in% BIOLOGICAL_PROCESS_GOs, .(OR_DOWN)]
        pval_down = facs[[.x]]@ORA$GO[Value %in% BIOLOGICAL_PROCESS_GOs, .(pval_DOWN)]
        
        g_OR_DIFF = (
            ggplot(or_diff, aes(OR_DIFF))
            + geom_histogram(binwidth=1)
            + geom_vline(aes(xintercept=10, colour='red'))
            + ggtitle(paste0(.x, ' OR_DIFF'))
        )
        g_pval_DIFF = (
            ggplot(pval_diff, aes(pval_DIFF))
            + geom_histogram(binwidth=0.01)
            + geom_vline(aes(xintercept=0.05, colour='red'))
            + ggtitle(paste0(.x, ' pval_DIFF'))
        )
        
        g_OR_UP = (
            ggplot(or_up, aes(OR_UP))
            + geom_histogram(binwidth=1)
            + geom_vline(aes(xintercept=10, colour='red'))
            + ggtitle(paste0(.x, ' OR_UP'))
        )
        g_pval_UP = (
            ggplot(pval_up, aes(pval_UP))
            + geom_histogram(binwidth=0.01)
            + geom_vline(aes(xintercept=0.05, colour='red'))
            + ggtitle(paste0(.x, ' pval_UP'))
        )
        
        g_OR_DOWN = (
            ggplot(or_down, aes(OR_DOWN))
            + geom_histogram(binwidth=1)
            + geom_vline(aes(xintercept=10, colour='red'))
            + ggtitle(paste0(.x, ' OR_DOWN'))
        )
        g_pval_DOWN = (
            ggplot(pval_down, aes(pval_DOWN))
            + geom_histogram(binwidth=0.01)
            + geom_vline(aes(xintercept=0.05, colour='red'))
            + ggtitle(paste0(.x, ' pval_DOWN'))
        )
        
        ggarrange(g_OR_DIFF, g_pval_DIFF,
                  g_OR_UP, g_pval_UP,
                  g_OR_DOWN, g_pval_DOWN,
                  nrow=3, ncol=2)
    }
)

# OR_DIFF distribution in a tissue
# or_diff = facs$Heart@ORA$GO[Value %in% BIOLOGICAL_PROCESS_GOs, .(OR_DIFF)]
# pval_diff = facs$Heart@ORA$GO[Value %in% BIOLOGICAL_PROCESS_GOs, .(pval_DIFF)]
# g1 = (
#     ggplot(or_diff, aes(OR_DIFF))
#     + geom_histogram(binwidth=1)
#     + ggtitle('Heart OR_DIFF')
# )
# 
# g2 = (
#     ggplot(pval_diff, aes(pval_DIFF))
#     + geom_histogram(binwidth=0.01)
#     + ggtitle('Heart pval_DIFF')
# )
# 
# ggarrange(g1, g2, ncol=2)
```


```{r}
cci_bp_gos = cci_gos[cci_gos %in% BIOLOGICAL_PROCESS_GOs]

significant_gos = function(ora_go, or_threshold) {
    
    sign_gos = ora_go[
        (OR_UP > or_threshold & pval_UP < 0.05)
        | (OR_DOWN > or_threshold & pval_DOWN < 0.05),
        # || (OR_DIFF > or_threshold && pval_DIFF < 0.05),
        Value
    ]
    # print(or_threshold)
    # print(dim(ora_go[Value==t, .(OR_UP, OR_DOWN)]))
    return(sign_gos)

}

significant_gos_from_scDiffCom = function(scDC, or_threshold) {
    ora_go = scDC@ORA$GO
    return(significant_gos(ora_go, or_threshold))
}

filter_gos = function(gos, subset, BP_GOs=BIOLOGICAL_PROCESS_GOs) {
    # GO_SUBSETS = c('BP', 'MF', 'CC')
    # if( !(subset %in% GO_SUBSETS) ) {
    #     stop("filter_gos: invalid argument `subset` ")
    # }
    if( subset=='BP' ) {
        return(
            gos[gos %in% BP_GOs]
        )
    } else if( subset=='MF' ) {
        stop('Not implemented')
    } else if( subset=='CC') {
        stop('Not implemented')
    } else {
        stop("filter_gos: invalid argument `subset`")
    }
}

# Will approximate
get_term_num_ancestors = function(go, terms) {
    return(
        map_int(
            terms,
            ~ (length(get_ancestors(go, .x)))
        ) - 1
    )
}

plot_hist = function(vec, xlab, bw=0.01, zoom=FALSE) {
    if(!zoom) {
        return(
            ggplot(data.frame(vec), aes(vec))
            + geom_histogram(binwidth=bw)
            + xlab(xlab)
            + theme_bw()
        )
    } else {
        return(
            ggplot(data.frame(vec), aes(vec))
            + geom_histogram(binwidth=bw)
            + xlab(xlab)
            + coord_cartesian(xlim=c(0, 0.05), ylim=c(0, 100))
            + theme_bw()
        )
    }
}

plot_scatter = function(data, zoom=FALSE) {
    if (!zoom) {
        return(
            ggplot(data, aes(x=NUM_ANCESTORS, y=TERM_FREQ))
            + geom_jitter()
            + geom_hline(yintercept=0.01)
            + geom_hline(yintercept=0.025)
            + geom_hline(yintercept=0.05)
            + theme_bw()
        )
    }
    else {
        return(
            ggplot(data, aes(x=NUM_ANCESTORS, y=TERM_FREQ))
            + geom_jitter()
            + coord_cartesian(xlim=c(0, 20), ylim=c(0, 0.10))
            + geom_hline(yintercept=0.01)
            + geom_hline(yintercept=0.025)
            + geom_hline(yintercept=0.05)
            + theme_bw()
        )
    }
    
}


tissues = names(facs)
map(
    tissues,
    function(.x) {

        gos = filter_gos(significant_gos_from_scDiffCom(facs[[.x]], 2), 'BP')
        freqs = get_term_frequencies(go, gos)
        freqs = freqs[ names(freqs) %in% gos ]  # for some reason increases
        num_ancestors = get_term_num_ancestors(go, gos)
        
        # length(gos)
        # length(freqs)
        # length(num_ancestors)
        
        freq_hist = plot_hist(freqs, 'Term frequency', 0.001, zoom=TRUE)
        ancest_hist = plot_hist(num_ancestors, 'Num ancestors', 1)
        
        data = data.table(
            GO=names(freqs), 
            TERM_FREQ=freqs, 
            NUM_ANCESTORS=num_ancestors
        )
        
        unzoom_scatter = plot_scatter(data, zoom=FALSE)
        zoomed_scatter = plot_scatter(data, zoom=TRUE)

        g = ggarrange(
            freq_hist, ancest_hist,
            unzoom_scatter, zoomed_scatter,
            nrow=2, ncol=2)
        return(
            annotate_figure(g, top=.x)
        )

    }
)

# Minimal set
# map(
#     tissues,
#     function(.x) {
#         gos = filter_gos(significant_gos_from_scDiffCom(facs[[.x]]), 'BP')
#         minim = minimal_set(go, gos)
#         # return(length(gos) - length(minim))
#         return(length(minim))
#     }
# )

# Consider information content
# get_term_info_content(go, filter_gos(significant_gos(facs$Lung@ORA$GO), 'BP'))
# minimal_set()
```



```{r}
# Get GO BP terms from all tissues with OR_DIFF > 10
tissues = names(facs)

# go_bp_significant_per_tissue = map(
#     tissues,
#     function(.x) {
#         GOs = facs[[.x]]@ORA$GO[
#             Value %in% BIOLOGICAL_PROCESS_GOs
#             && (
#                 (OR_UP > 10 & pval_UP < 0.05)
#                 | (OR_DOWN > 10 & pval_DOWN < 0.05)
#                 | (OR_DIFF > 10 & pval_DIFF < 0.05)
#             ),
#             Value
#         ]
#         return(GOs)
#     }
# )

go_bp_significant_per_tissue = map(
    tissues,
    function(.x) {
        return(significant_gos(facs[[.x]]@ORA$GO, 2))
    }
)

# # Num significant per tissue
# map_int(
#     go_bp_significant_per_tissue,
#     length
# )
# 
# go_bp_sign_tissue_union = reduce(
#     go_bp_significant_per_tissue,
#     union
# )
# go_bp_sign_tissue_intersection = reduce(
#     go_bp_significant_per_tissue,
#     intersect
# # )
# message('Significant GO BP tissue union: ', length(go_bp_sign_tissue_union))
# message('Significant GO BP tissue intersection: ', length(go_bp_sign_tissue_intersection))

```

## Augment ORA GO tables with term frequency and num_ancestors for sorting 

```{r}
two_point_interval_label = function(
    values, low_th, up_th, 
    low_lab, mid_lab, high_lab) {
    
    return(
        ifelse(
            (values < low_th), low_lab, ifelse(
                (values > up_th), high_lab, mid_lab
            )
        )
    )
}
# dataset=facs[['Liver']]
# or_threshold = 2
# subset = 'BP'

compute_reduced_ora_go = function(dataset, or_threshold=2, subset='BP', gene_ontology=go) {
    
    gos = filter_gos(significant_gos_from_scDiffCom(dataset, or_threshold), subset)
    freqs = get_term_frequencies(gene_ontology, gos)
    freqs = freqs[ names(freqs) %in% gos ]  # for some reason increases
    num_ancestors = get_term_num_ancestors(gene_ontology, gos)
    # info_content = get_term_info_content(gene_ontology, gos)
    # info_content = info_content[ names(info_content) %in% gos ]  # for some reason increases

    if( !(  (length(gos) == length(freqs))
            & (length(gos) == length(num_ancestors))
            # & (length(gos) == length(info_content))
           )) {
            print(length(gos))
            print(length(freqs))
            print(length(num_ancestors))
            # print(length(info_content))
        stop("gos, freqs, num_ancestors and info_content must have same len.")    
    }
    
    names(freqs) = gos
    names(num_ancestors) = gos
    # names(info_content) = gos
    
    dt = copy(dataset@ORA$GO)
    
    dt[Value %in% gos, c('Num_ancestors', 'Term_frequency') := list(
        num_ancestors[Value],
        freqs[Value]
        # info_content[Value]
    )]

    dt_s = dt[Value %in% gos]  # subset the relevant GO

    
    FREQUENT_LABEL = 'frequent'
    RARE_LABEL = 'rare'
    UNCATEGORIZED_LABEL = 'uncategorized'
    lower_th_freq = quantile(freqs, 0.2)
    upper_th_freq = quantile(freqs, 0.8)
    frequency_label = two_point_interval_label(
        freqs, lower_th_freq, upper_th_freq, 
        RARE_LABEL, UNCATEGORIZED_LABEL, FREQUENT_LABEL
    )
    
    GENERAL_LABEL = 'general'
    SPECIFIC_LABEL = 'specific'
    lower_th_num_ancestors = quantile(num_ancestors, 0.2)
    upper_th_num_ancestors = quantile(num_ancestors, 0.8)
    generality_label = two_point_interval_label(
        num_ancestors, lower_th_num_ancestors, upper_th_num_ancestors, 
        GENERAL_LABEL, UNCATEGORIZED_LABEL, SPECIFIC_LABEL
    )
    
    # print(length(gos))
    # print(length(frequency_label))
    # print(length(generality_label))
    # print(dim(dt_s))
    # print(sum(dt[,Value] %in% gos))

    dt_s[, c('Frequency', 'Generality') := .(
        frequency_label, generality_label
    )]
    
    dt_s = dt_s[order(-Term_frequency, Num_ancestors), .(Value, OR_UP, OR_DOWN, Term_frequency, Num_ancestors, Frequency, Generality)]
    dt_s = dt_s[ (Frequency %in% (FREQUENT_LABEL)) & (!(Generality %in% (UNCATEGORIZED_LABEL)))]
    dt_s[, Biological_process := Term(Value)]
    dt_s = dt_s[, .(
        Value,
        Biological_process,
        OR_UP, OR_DOWN,
        Frequency, Generality,
        Term_frequency, Num_ancestors
    )]
    return(dt_s)
}

dt = facs[['Liver']]@ORA$GO
dt_s = compute_reduced_ora_go(facs[['Liver']], 2, 'BP')
dt_s
write.csv(dt_s, file='../../liver_go.csv')

dt[, list(Value, OR_UP, OR_DOWN)]

dts = map(
    tissues,
    ~ compute_reduced_ora_go(facs[[.x]], 2, 'BP')
)
```


## FPM for GO terms

    Another approach to simplify GO terms and extract insight into biological processes is to perform frequent pattern mining (FPM) on the GO terms.
    The setup for FPM is as follows. In a per-tissue LR-interaction filtered data.table rows represent LR-interactions. We assign a set of GO terms to each LR-interaction based on aggregating GO terms from the ligand and receptor. We can view it as a frequent pattern mining problem: every LR-interaction "buys" a "bag" of GO terms + p-values + logFC. We then perform FPM to identify to identify conditional rules (that model implication relationships) of the following form:
        
    `    {Set of GO terms of variable size} => {p-val=significant, Direction=UP/DOWN}`

    We evaluate every rule by Fisher test.


```{r}
# library(scDiffCom)
# source('../src/fpm.R')
# 
# data("LR6db")
# 
# LR_GOs = scDiffCom::LR6db$LR6db_GO$LR_GO

# Merge to create transaction data structure

# Apply FPM

# Evaluate FPM with metrics

# Select the significant ones

# Characterize GO terms sets by depth/term frequencies (or sort the set by these)



```
