---
title: "R Notebook"
output: html_notebook
---

We'll look at the cell ontology.

General purpose ontology manipulation in R can be done with ontologyX tools (ontologyIndex, ontologyPlot, ontologySimilarity?) and ontoProc offers bio-geared ontology functionality.

```{r}
library(ontoProc)
library(ontologyIndex)
library(ontologyPlot)
```

General ontologies are often stored in .owl format and, in terms of technologies, ontologies can be served by graph databases. OBO, the open biological ontologies foundation, offers bio-related ontologies in .OBO format:
http://www.obofoundry.org/ontology/cl.html

getCellOnto() can be used to extract ontology and represent it as orthology_index class as defined by the orthologyIndex package. orthologyIndex is a package for general-purpose low-level ontology manipulation.

Let's explore oncoProc functionality
```{r}
cl = getCellOnto()

# Plot CL for terms
cell_v = c("CL:0000492", "CL:0001054", "CL:0000236", "CL:0000625",
   "CL:0000576", "CL:0000623", "CL:0000451", "CL:0000556")
onto_plot2(cl, cell_v)

# feats = CLfeats(cl, "CL:0000492")

# Names from cell line ontology
# cleannames = cleanCLOnames()

# Start app and dataframes with cell types properties
# ctmarks(cl)

# onto_roots(cl)
recognizedPredicates()
```

Explore ontologyIndex functionality
```{r}
cl = getCellOnto()  # already ontology_index class

# Explore the data structure
str(cl, max.level=1)
term = "CL:0000492"
cl$name[term]
cl$parents[term]
cl$name[cl$ancestors[[term]]]

# Get term properties
get_term_property(
    ontology=cl, 
    property="ancestors",
    term="CL:0000492",
    as_names=TRUE
)

# Remove redundant terms from a set of terms
terms = c("CL:0000492", "CL:0001054", "CL:0000236", "CL:0000625",
   "CL:0000576", "CL:0000623")
minimal = minimal_set(cl, terms)

# Find all ancestors of a set of terms
ancestors_ids = get_ancestors(cl, terms)
cl$name[ancestors_ids]

# Operations on subclasses
# intersection_with_descendants()
# exclude_descendants()
# prune_descendants()

properties = names(cl)
# cl$has_high_plasma_membrane_amount

```

Let's use the tools to explore some cell ontologies. We can see that 'bladder cell' is a term identifying all types of cells identified in the bladder.

```{r}
# Let's look into the bladder cells
bladder_ids = c('CL:1001319','CL:1001428','CL:0000738')
cl$name[bladder_ids]
onto_plot2(cl, bladder_ids)

# Let's look into all descendants of the bladder cell
bladder_descendants = get_descendants(cl, 'CL:1001319')
onto_plot2(cl, bladder_descendants)

macrophage_descendants = get_descendants(cl, cl$id[cl$name == 'macrophage'])
onto_plot2(cl, macrophage_descendants)

# Cellname mapping to cell ontology ----
# Let's explore the brain's cell types
brain_celltypes_names = c(
    'Bergmann glial cell',
    'CD8-positive, alpha-beta T cell',
    'T cell',
    'astrocyte',
    'brain pericyte',
    'endothelial cell',
    'ependymal cell',
    'interneuron',
    'macrophage',
    'mature NK T cell',
    'medium spiny neuron',
    'microglial cell',
    'neuroepithelial cell',
    'neuron',
    'neuronal stem cell',
    'oligodendrocyte',
    'oligodendrocyte precursor cell'
)

cellnames = cl$name

library(purrr)
map_lgl(
    brain_celltypes_names,
    ~ map_cellname(.x, cellnames)
)

map_cellname(brain_celltypes_names[6], cellnames)

map_cellname = function (name, cellnames){
    idxs = grep(
        pattern=name,
        x=cellnames, 
        ignore.case=TRUE
    )
    return(length(idxs)>0)
}
```


We can use cell ontology and the above functionality to help annotate cell types in a couple of ways:
- for each cell ontology id, get the path to root for manual inspection
- for each cell ontology id, get the properties
- for each tissue, get the minimal set of terms. Use these to evaluate how much the given tissue covers the space of cell types (e.g. width of the ontology subgraph obtained by subsetting the general ontology by the terms in the minimal set).
- for each tissue, plot the relevant subgraph and the non-redundant, i.e. minimal, subgraph. 


```{r}

celltype_names = brain_celltypes_names
ontology = cl

search_ontology_cellnames = function(ontology, cellnames) {
    
    ontology_cellnames = ontology$name
    
    search_cellname = function(cellnames_to_search, cellname_to_search) {
        idxs = grep(
            pattern=cellname_to_search,
            x=cellnames_to_search, 
            ignore.case=TRUE
        )
        is_identified = length(idxs) > 0
        return(is_identified)
    }
    
    identified_mask = map_lgl(
        cellnames,
        ~ search_cellname(ontology_cellnames, .x)
    )
    unidentified_mask = !identified_mask
    
    message(paste0("search_ontology_cellnames: ", 
                   sum(unidentified_mask), 
                   ' cell names not found in ontology')
    )
    
    return(cellnames[unidentified_mask])
}
```



```{r}
id = "CL:0000492"  # a cell ontology id
# cell ontology unique ids corresponding to cell types from brain
ids = c("CL:0000644", "CL:0000127", "CL:0000128",
        "CL:2000043", "CL:0000115", "CL:0000540", 
        "CL:0002453")
cl = getCellOnto()  # cell ontology

get_path_to_root = function(ontology=cl,
                            id) {
    stop('Not implemented')
}

get_properties = function(ontology=cl,
                          id) {
    stop('Not implemented')
}

compute_minimal_set = function(ontology=cl,
                               ids) {
    stop('Not implemented')
}

plot_ontology_subset = function(ontology=cl,
                                ids) {
    stop('Not implemented')
}
```

In the TMS paper, they do graph-based clustering on PCA-reduced representation of raw data. They then map cell types to <CL / previous paper data>: if 1-1 ok, else flag for manual/expert annotation by the tissue expert. That explains the inconsistency between cell ontology and final cell labels.

A solution could be to use the current version of cell ontology. Since some time has passed, it may be the case that the new cell types have been resolved in CL. Maybe they're collaborating with CL to resolve. We should flag ourselves the cell types that can't be resolved by current CL and discuss with the TMS / CL team.
