---
title: "R Notebook"
---

This notebook presents the heatmaps and bipartite graphs.

```{r}
library(data.table)
library(purrr)
library(pheatmap)
library(RColorBrewer)
library(scales)
library(igraph)


#RESULTS_PATH = '../../data_scAgeCom/testing/scdiffcom_tms_facs_mixed_Spleen.rds'
#results = readRDS(RESULTS_PATH)
results <- readRDS("../../data_scAgeCom/analysis/a4_data_results_nlog.rds")
results <- results$droplet_nlog$Limb_Muscle

results$ORA$OR_UP[is.infinite(results$ORA$OR_UP)] <- max(results$ORA$OR_UP[is.finite(results$ORA$OR_UP)])
results$ORA$OR_DOWN[is.infinite(results$ORA$OR_DOWN)] <- max(results$ORA$OR_DOWN[is.finite(results$ORA$OR_DOWN)])
results$ORA$OR_DIFF[is.infinite(results$ORA$OR_DIFF)] <- max(results$ORA$OR_DIFF[is.finite(results$ORA$OR_DIFF)])

results$ORA$OR_UP[results$ORA$OR_UP == 0] <- min(results$ORA$OR_UP[results$ORA$OR_UP > 0])
results$ORA$OR_DOWN[results$ORA$OR_DOWN == 0] <- min(results$ORA$OR_DOWN[results$ORA$OR_DOWN > 0])
results$ORA$OR_DIFF[results$ORA$OR_DIFF == 0] <- min(results$ORA$OR_DIFF[results$ORA$OR_DIFF > 0])
source('../src/utils.R')
results = add_dummy_tissue(results)

str(results, max.level = 1)
```

# Heatmaps

Internally, we convert dt_ora to dt_ctypes, a data.table that storesedge data - a row describes an edge between a ligand and receptor celltype with edge attributes including: OR, OR_UP, OR_DOWN. This data.table is used to build the heatmaps.

Below code was initially made to work with multiple datasets and tissues. For now we test on a single tissue, but code is left  for future development.

```{r}
source('../src/visualizations.R')

create_heatmaps_dirs = function(base_dir, dir_names){
  # Create paths
  heatmaps_dirs = map_chr(
    dir_names,
    ~ file.path(base_dir, .x)
  )  
  # Create dirs at paths
  map(
    heatmaps_dirs,
    ~ dir.create(.x, showWarnings=FALSE)
  )
  heatmaps_dirs = as.list(heatmaps_dirs)
  names(heatmaps_dirs) = dir_names
  return(heatmaps_dirs)
}

construct_heatmaps = function(ora_dt, dir){
  dt_ctypes = get_celltypes_enrichment(ora_dt)
  build_heatmaps(dt_ctypes, dir)
}

HEATMAPS_BASE_DIR = "../../data_scAgeCom/analysis/heatmaps"
heatmaps_dirs = create_heatmaps_dirs(HEATMAPS_BASE_DIR, c('test_dataset'))
construct_heatmaps(results$ORA, heatmaps_dirs$test_dataset)
```

# Number of cells bias

```{r}
library(ggplot2)
dt = results$scdiffcom_dt_filtered

(ggplot(data=dt, aes(L_NCELLS_YOUNG, R_NCELLS_YOUNG))
 + geom_point(aes(color=DIFFERENTIALLY_EXPRESSED))
 + ggtitle('N cell bias: young')
)

(ggplot(data=dt, aes(L_NCELLS_OLD, R_NCELLS_OLD))
 + geom_point(aes(color=DIFFERENTIALLY_EXPRESSED))
 + ggtitle('N cell bias: old')
)

(ggplot(data=dt, aes(L_NCELLS_YOUNG, L_NCELLS_OLD))
 + geom_point(aes(color=DIFFERENTIALLY_EXPRESSED))
 + ggtitle('N cell bias: ligand cells young vs old')
)

(ggplot(data=dt, aes(R_NCELLS_YOUNG, R_NCELLS_OLD))
 + geom_point(aes(color=DIFFERENTIALLY_EXPRESSED))
 + ggtitle('N cell bias: receptor cells young vs old')
)
# coord_cartesian(xlim = c(0, 5))
```

# Bipartite graphs for cell-cell communication


```{r}
source("../src/visualizations.R")
source("../src/utils.R")  # for loading GRAPH_CONFIG

dir = "../../data_scAgeCom/analysis/Graphs"
if ( !dir.exists(dir) ) {dir.create(dir)}

analyze_Graph(
  results$ORA,
  results$scdiffcom_dt_filtered,
  "DummyTissue",
  config=GRAPH_CONFIG,
  use_adjpval=TRUE,
  disperse=TRUE,
  dir=dir,
  analysis_name="DummyTest")


### Below code was being used for multiple datasets/tissues analyses
# analyze <- function(dataset_name, use_adjpval=FALSE) {
#   tissues = unique(results$ora[[dataset_name]]$Tissue)
#   
#   # rm All
#   tissues = tissues[ !(tissues == 'All') ]
#   
#   if( use_adjpval) {
#     analysis_name = paste0(dataset_name, '_with_adjpval')
#   } else {
#     analysis_name = paste0(dataset_name, '_without_adjpval') 
#   }
# 
#   map(
#     tissues, 
#     ~ analyze_Graph(
#       results$ora[[dataset_name]],
#       results$filtered[[dataset_name]]$results,
#       .x,
#       config=GRAPH_CONFIG,
#       use_adjpval=use_adjpval,
#       dir=dir, 
#       analysis_name=analysis_name
#       )
#     )
# }
# 
# 
# map(
#   names(results$filtered),
#   ~ analyze(.x, use_adjpval=TRUE)
# )
```


