---
title: "R Notebook"
---

Celltypes analysis not integrated yet with multiple subunits ligand-receptor complexes.

Get the datasets.

```{r}
DATA_PATH = "../../../data/pipeline_output/analysis_4_data_diffcom_filter.rds"
datasets = readRDS(DATA_PATH)

message("Dataset composition:")
str(datasets, max.level=1)

message("Dataset structure:")
str(datasets[[1]], max.level=1)

message("Dataset tissues:")
map(datasets, ~ unique(.x[["TISSUE"]]))


cols = get_default_colnames()

datasets = map(
    datasets,
    ~ analyze_CCI(
      data = .x,
      cols = cols,
      cutoff_score = quantile(c(.x$LR_SCORE_old, .x$LR_SCORE_young), 0.75)
    )
)
```

# Celltypes
Extract unique celltypes from tms facs and droplet.

```{r}

extract_unique_celltypes <- function(dt) {
    cells = dt[,
               .('Celltype'=union(L_CELLTYPE, R_CELLTYPE)),
               by=TISSUE]
    return(cells)
}

celltypes = map(
    datasets, 
    extract_unique_celltypes
)

celltypes_all = reduce(
    celltypes,
    ~ funion(.x, .y)
)

dataset_has_celltype <- function(dt_name, celltype, tissue) {
    return(
        celltype %in% celltypes[[dt_name]][TISSUE==tissue, Celltype]
    )
}

vec_dataset_has_celltype = Vectorize(dataset_has_celltype, 
                                     vectorize.args = c('celltype', 'tissue')
                                     )

vec_dataset_has_celltype('calico', 'macrophage', 'Kidney')

celltypes_all[, names(datasets) := list(
    vec_dataset_has_celltype('calico', Celltype, TISSUE),
    vec_dataset_has_celltype('calico_sub', Celltype, TISSUE),
    vec_dataset_has_celltype('tms_facs', Celltype, TISSUE),
    vec_dataset_has_celltype('tms_droplet', Celltype, TISSUE)
)]

### ADD CELLTYPE and GENE in addition to L_GENE, L_CELLTYPE, ...

write.table(celltypes_all, '../../celltypes.csv', sep='\t', quote = FALSE, row.names = FALSE)
celltypes_all
```

# Cell ontology

We'll look at the cell ontology.

General purpose ontology manipulation in R can be done with ontologyX tools (ontologyIndex, ontologyPlot, ontologySimilarity?) and ontoProc offers bio-geared ontology functionality.

```{r}
library(ontoProc)
library(ontologyIndex)
library(ontologyPlot)
library(purrr)
```

General ontologies are often stored in .owl format and, in terms of technologies, ontologies can be served by graph databases. OBO, the open biological ontologies foundation, offers bio-related ontologies in .OBO format:
http://www.obofoundry.org/ontology/cl.html

getCellOnto() can be used to extract ontology and represent it as orthology_index class as defined by the orthologyIndex package. orthologyIndex is a package for general-purpose low-level ontology manipulation.

Let's explore oncoProc functionality
```{r}
# cl = getCellOnto()  # gets an old version!
# cl = get_ontology(file='../../data/cl.obo')


cl$name['CL:0000127']

# Plot CL for terms
cell_v = c("CL:0000492", "CL:0001054", "CL:0000236", "CL:0000625",
   "CL:0000576", "CL:0000623", "CL:0000451", "CL:0000556")
onto_plot2(cl, cell_v)

# feats = CLfeats(cl, "CL:0000492")

# Names from cell line ontology
# cleannames = cleanCLOnames()

# Start app and dataframes with cell types properties
# ctmarks(cl)


# onto_roots(cl)
recognizedPredicates()
```

Explore ontologyIndex functionality
```{r}
# Explore the data structure
str(cl, max.level=1)
term = "CL:0000492"
cl$name[term]
cl$parents[term]
cl$name[cl$ancestors[[term]]]

# Get term properties
get_term_property(
    ontology=cl, 
    property="ancestors",
    term="CL:0000492",
    as_names=TRUE
)

# Remove redundant terms from a set of terms
terms = c("CL:0000492", "CL:0001054", "CL:0000236", "CL:0000625",
   "CL:0000576", "CL:0000623")
minimal = minimal_set(cl, terms)

# Find all ancestors of a set of terms
ancestors_ids = get_ancestors(cl, terms)
cl$name[ancestors_ids]

# Operations on subclasses
# intersection_with_descendants()
# exclude_descendants()
# prune_descendants()

properties = names(cl)
# cl$has_high_plasma_membrane_amount

```

Let's use the tools to explore some cell ontologies. We can see that 'bladder cell' is a term identifying all types of cells identified in the bladder.

```{r}
# Let's look into the bladder cells
bladder_ids = c('CL:1001319','CL:1001428','CL:0000738')
cl$name[bladder_ids]
onto_plot2(cl, bladder_ids)

# Let's look into all descendants of the bladder cell
bladder_descendants = get_descendants(cl, 'CL:1001319')
onto_plot2(cl, bladder_descendants)

macrophage_descendants = get_descendants(cl, cl$id[cl$name == 'macrophage'])
onto_plot2(cl, macrophage_descendants)

# Cellname mapping to cell ontology ----
# Let's explore the brain's cell types
brain_celltypes_names = c(
    'Bergmann glial cell',
    'CD8-positive, alpha-beta T cell',
    'T cell',
    'astrocyte',
    'brain pericyte',
    'endothelial cell',
    'ependymal cell',
    'interneuron',
    'macrophage',
    'mature NK T cell',
    'medium spiny neuron',
    'microglial cell',
    'neuroepithelial cell',
    'neuron',
    'neuronal stem cell',
    'oligodendrocyte',
    'oligodendrocyte precursor cell'
)

cellnames = cl$name
map_cellname = function (name, cellnames){
    idxs = grep(
        pattern=name,
        x=cellnames, 
        ignore.case=TRUE
    )
    return(length(idxs)>0)
}
map_lgl(
    brain_celltypes_names,
    ~ map_cellname(.x, cellnames)
)

map_cellname(brain_celltypes_names[6], cellnames)


```


We can use cell ontology and the above functionality to help annotate cell types in a couple of ways:
- for each cell ontology id, get the path to root for manual inspection
- for each cell ontology id, get the properties
- for each tissue, get the minimal set of terms. Use these to evaluate how much the given tissue covers the space of cell types (e.g. width of the ontology subgraph obtained by subsetting the general ontology by the terms in the minimal set).
- for each tissue, plot the relevant subgraph and the non-redundant, i.e. minimal, subgraph. 


```{r}

celltype_names = brain_celltypes_names
ontology = cl

###

build_search_pattern = function(cellname) {
    # > grep('^T cell| T cell$| T cell |^T cell$', c('neurat cell', 'T cell', 'alpha T cell'), ignore.case=TRUE)
    pattern = paste0(
        '^', cellname,
        '|', ' ', cellname, '$',
        '|', ' ', cellname, ' ',
        '|', '^', cellname, '$'
    )
    return(pattern)
}

remove_obsolete_cellnames = function(cellnames) {
    remove_mask = grepl(
        pattern='obsolete',
        x=cellnames,
        ignore.case=TRUE
    )
    return(cellnames[!remove_mask])
}

### Search ontology for the cellnames
ids_from_cellname = function(ontolgy, cellname_to_search) {
    
    if (length(cellname_to_search) > 1) {
        stop('find_cellname_ids: input cellname of length > 1')
    }
    
    pattern = build_search_pattern(cellname_to_search)
    
    
    # Just extracting names extracts GO and other terms which are not cellnames,
    #  but which might include, e.g. GO: T-cell cytotoxicity.
    # ontology_cellnames = ontology$name
    
    # ontology_cellnames = remove_noncellnames_terms(ontology$name)
    CL_terms_mask = map_lgl(
        names(cl$name),
        ~ strsplit(.x, ':')[[1]][1] == 'CL'
    )
    ontology_cellnames = cl$name[CL_terms_mask]
    ontology_cellnames = remove_obsolete_cellnames(ontology_cellnames)
    
    
    indexes = grep(
        pattern=pattern,
        x=ontology_cellnames, 
        ignore.case=TRUE
    )
    
    # Indexes in ontology_cellnames != indexes in ontology$name, since
    #  ontology_cellnames are a subset that represent only cell names.
    #  This is the reason for the extra lines below.
    found_cellnames = ontology_cellnames[indexes]
    mask = ontology$name %in% found_cellnames
    ids = ontology$id[mask]
    return(ids)
}

cellname_exists = function(ontology, cellname_to_search) {
    ids = ids_from_cellname(ontology, cellname_to_search)
    is_found = length(ids) > 0
    return(is_found)
}

find_existing_cellnames = function(ontology, cellnames) {
    identified_mask = map_lgl(
        cellnames,
        ~ cellname_exists(ontology, .x)
    )
    
    unidentified_mask = !identified_mask
    
    return(
        list(
            found=cellnames[identified_mask],
            notfound=cellnames[unidentified_mask]
        )
    )
}

### Get the common ancestor for cellname hits
compute_common_ancestor_mask = function(cl, ids){
    desc_matrix = get_term_descendancy_matrix(cl, terms=ids)
    common_ancestor_mask = rowSums(desc_matrix) == (length(ids) - 1)
    if (sum(common_ancestor_mask) > 1) {
        stop('compute_common_ancestor_mask: mask sum is > 1')
    }
    return(common_ancestor_mask)
}

common_ancestor_exists = function(cl, ids) {
    mask = compute_common_ancestor_mask(cl, ids)
    exists = sum(mask) == 1
    return(exists)
}

get_ancestor = function(ontology, ids, cellname) {
    if( !common_ancestor_exists(ontology, ids) ){
        
        ancestor_candidate = get_ancestor_candidate(ontology, ids)
        non_descendants = get_nondescendants(cl, ancestor_candidate, ids)
        
        message('get_common_ancestor: Common ancestor does not exist for ',
                cellname, '. Ancestor inferred as term with most descendants.',
                ' Inferred ancestor = ', 
                paste(ontology$name[ancestor_candidate], sep='', collapse='; '),
                '; Non-descendants = ',
                paste(ontology$name[non_descendants], sep='', collapse='; '), 
                '\n'
        )
        
        return(ancestor_candidate)
    }
    mask = compute_common_ancestor_mask(ontology, ids)
    common_ancestor = ids[mask]
    return(common_ancestor)
}

### Resolve each cellname to a single id
resolve_cellname = function(ontology, cellname){
        ids = ids_from_cellname(ontology, cellname)
        common_ancestor_id = get_ancestor(ontology, ids, cellname)
        return(common_ancestor_id)
    }

get_cellnames_ids = function(ontology, cellnames) {
    # Check all cellnames are found in ontology
    all_cellnames_found = (
        length(find_existing_cellnames(ontology, cellnames)$notfound) == 0
    )
    
    if( !all_cellnames_found ){
        stop('get_cellnames_ids: Some cellnames are not found.')
    }
    
    resolved_ids = map_chr(
        cellnames,
        ~ resolve_cellname(ontology, .x)
    )
    
    return(resolved_ids)
}

###
get_topk_of_hierarchy_ids = function(cl, ids, k) {
    if( k>length(ids) ){
        stop('get_topk_of_hierarchy_ids: Can\'t plot k > num ids')
    }
    m = get_term_descendancy_matrix(cl, ids)
    num_descendants_in_ids = rowSums(m)
    descending_order_idxs = order(-num_descendants_in_ids)
    topk_ids = names(num_descendants_in_ids[descending_order_idxs][1:k])
    return(topk_ids)
}
plot_topk_of_hierarchy = function(cl, ids, k) {
    topk_ids = get_topk_of_hierarchy_ids(cl, ids, k)
    onto_plot2(cl, topk_ids)
}
get_ancestor_candidate = function(cl, ids){
    return(
        names(rowSums(get_term_descendancy_matrix(cl, ids)))[1]
    )
}
get_nondescendants = function(cl, root, terms) {
    return(exclude_descendants(cl, root, terms))
}


# Case 1: obsolete T cell cellnames
# ids = ids_from_cellname(cl, 'T cell')
# plot_topk_of_hierarchy(cl, ids, 40)

# Case 2: strange cases of non-descendants
ids = ids_from_cellname(cl, 'endothelial cell')
ancestor_candidate = get_ancestor_candidate(cl, ids)
non_descendants = get_nondescendants(cl, ancestor_candidate, ids)
print(cl$name[ancestor_candidate])
print(cl$name[non_descendants])
onto_plot2(cl, non_descendants)
get_cellnames_ids(cl, cellnames_found)

### Plots at high-level

analyze_celltypes = function(cl, cellnames, analysis_name) {
    existing_cellnames = find_existing_cellnames(cl, cellnames)
    cellnames_found = existing_cellnames$found
    
    ids = get_cellnames_ids(cl, cellnames_found)
    ids_minimal = minimal_set(cl, ids)
    
    BASE_PATH = '../../../results/Cellnames'
    if( !dir.exists(BASE_PATH)) {dir.create(BASE_PATH)}
    
    analysis_path = file.path(BASE_PATH, analysis_name)
    # if( !dir.exists(analysis_path)) {dir.create(BASE_PATH)}
    dir.create(analysis_path)
    
    BASE_WIDTH = 800
    BASE_HEIGHT = 800
    LARGE_WIDTH = 3000
    LARGE_HEIGHT = 3000
    POINTSIZE = 25
    
    plot_and_save = function(cl, ids, analysis_path, filename, width, height){
        filename = file.path(analysis_path, filename)
        png(filename, width=width, height=width, pointsize=POINTSIZE)
        onto_plot2(cl, ids)
        dev.off()
    }
    
    if( length(ids) > 1) {
        # filename = file.path(analysis_path, 'cellnames.png')
        # png(filename, width=width, height=width, pointsize=pointsize)
        # onto_plot2(cl, ids)
        # dev.off()
        plot_and_save(cl, ids, analysis_path, 'cellnames.png', BASE_WIDTH, BASE_HEIGHT)
        plot_and_save(cl, ids, analysis_path, 'cellnames_LARGE.png', LARGE_WIDTH, LARGE_HEIGHT)
    } else {
        message('analyze_celltypes: Insufficient cellnames for ',
                analysis_name)
    }
    
    if( length(ids_minimal) > 1) {
        # filename = file.path(analysis_path, 'cellnames_minimal_set.png')
        # png(filename, width=width, height=width, pointsize=pointsize)
        # onto_plot2(cl, ids_minimal)
        # dev.off()
        plot_and_save(cl, ids_minimal, analysis_path, 'cellnames_minimal.png', BASE_WIDTH, BASE_HEIGHT)
        plot_and_save(cl, ids_minimal, analysis_path, 'cellnames_minimal_LARGE.png', LARGE_WIDTH, LARGE_HEIGHT)
    } else {
        message('analyze_celltypes: Insufficient cellnames in minimal set for ',
                analysis_name)
    }
}

# analyze_celltypes(cl, brain_celltypes_names, 'brain_test')

###
# props = get_ids_properties(ids)  # extract properties for each cell and store.
# onto_plot2(cl, ids)

# Coloring terms based on properties
# termset = list(
#     A = ids[1:3],
#     B = ids[4:length(ids)]
# )
# onto_plot(cl, term_sets = termset, fillcolor = colour_by_term_set)

###
existing_cellnames = find_existing_cellnames(cl, brain_celltypes_names)
cellnames_found = existing_cellnames$found
cellnames_not_found = existing_cellnames$notfound

message("?: Cellnames not found in ontology: ", 
        paste(cellnames_not_found, sep='', collapse='; '), '\n')

ids = get_cellnames_ids(cl, cellnames_found)
message('Resolved cellnames: ', paste(cl$name[ids], sep='', collapse='; '), '\n')

# {
# png('test.png', width=800, height=800, pointsize=20)
# onto_plot2(cl, ids)
# dev.off()
# }

# {
# onto_plot2(cl, ids)
# minimal = minimal_set(cl, ids)
# onto_plot2(cl, minimal)
# }

# cellname_ids = find_cellname_ids(cl, cellnames_found)

```

Let's get graphs for cell ontology subset for a given tissue and the corresponding minimal set.

```{r}
library(data.table)

# Get cellnames for every tissue
CELLTYPES_PATH = '../../../data/cell_types_with_counts - cell_types_with_counts.tsv'
dt_ct = data.table(read.csv(CELLTYPES_PATH, sep='\t'))

# Indicate if cell exists in each dataset
dt_ct[, is_in_callico := !(is.na(Calico.old.male) & is.na(Calico.young.male))]
dt_ct[, is_in_droplet := !(is.na(Droplet.old.female) & is.na(Droplet.old.male)
                        & is.na(Droplet.young.female) & is.na(Droplet.young.male))]
dt_ct[, is_in_facs := !(is.na(FACS.old.female) & is.na(FACS.old.male)
                        & is.na(FACS.young.female) & is.na(FACS.young.male))]

# Add cellname column
dt_ct[, Cellname := .(map_chr(
    dt_ct[, Cell.type],
    ~ strsplit(.x, '_')[[1]][length(strsplit(.x, '_')[[1]])]
))]

get_tissue_cellnames = function(dt, tissue){
    return(
        dt[Tissue==tissue, Cellname]
    )
}

tissues = unique(dt_ct[, Tissue])
cellnames = map(
    tissues,
    ~ get_tissue_cellnames(dt_ct, .x)
)
names(cellnames) = tissues

# Perform analysis across all tissues
map(
    tissues,
    ~ analyze_celltypes(cl, cellnames[[.x]], .x)
)
```

```{r}




```


In the TMS paper, they do graph-based clustering on PCA-reduced representation of raw data. They then map cell types to <CL / previous paper data>: if 1-1 ok, else flag for manual/expert annotation by the tissue expert. That explains the inconsistency between cell ontology and final cell labels.

A solution could be to use the current version of cell ontology. Since some time has passed, it may be the case that the new cell types have been resolved in CL. Maybe they're collaborating with CL to resolve. We should flag ourselves the cell types that can't be resolved by current CL and discuss with the TMS / CL team.

