---
title: "R Notebook"
---


```{r}
# Detect changed column names
new = names(obj@cci_detected)
old = names(obj_old@cci_detected)

intersection = new[(new %in% old)]
new_excl = new[!(new %in% old)]
old_excl = old[!(old %in% new)]


head(obj_old@cci_detected[, ..old_excl])
head(obj@cci_detected[, intersection, with=FALSE])

# LR_SORTED -> encapsulated in `CCI`
# DIFFERENTIALLY_EXPRESSED -> IS_CCI_DE
# DIFFERENTIAL_DIRECTION -> DE_DIRECTION
# REGULATION_SIMPLE -> X
# CCI_DETECTED_AND_SIGNIFICANT_IN_YOUNG -> ?? IS_CCI_EXPRESSED_YOUNG, IS_CCI_SCORE_YOUNG, IS_CCI_SPECIFIC_YOUNG
# CCI_DETECTED_AND_SIGNIFICANT_IN_OLD -> ...
# CCI_DETECTED_YOUNG -> IS_CCI_DETECTED_YOUNG
# CCI_DETECTED_OLD -> ...
# EMITTER_NCELLS_YOUNG -> NCELLS_EMITTER_YOUNG
# EMITTER_NCELLS_OLD -> ...
# L1_DETECTED_YOUNG -> L1_DETECTION_RATE_YOUNG
# 

```
```{r}
head(obj@cci_detected[, ..new_excl])
```


```{r}
cci_detected = obj@cci_detected
head(cci_detected)
```

```{r}
lr_sorted = 'Abca1_Calm1'
```


```{r}
library(scDiffCom)
# library(ComplexUpset)

lrdb = scDiffCom::LRdb_mouse$LRdb_curated
head(lrdb)
```

```{r}
lr_cols = c('LIGAND_1', 'RECEPTOR_1', 'LIGAND_2', 'RECEPTOR_2', 'RECEPTOR_3')
# lrdb[, ..lr_cols]

lrdb
lrdb_mono = lrdb[is.na(LIGAND_2) & is.na(RECEPTOR_2) & is.na(RECEPTOR_3)]
lrdb_poly = lrdb[!is.na(LIGAND_2) | !is.na(RECEPTOR_2) | !is.na(RECEPTOR_3)]


library(igraph)
library(purrr)
dim(lrdb[is.na(LIGAND_1) | is.na(RECEPTOR_1)])[1] == 0

make_test_mono_graph <- function() {
  t = igraph::make_empty_graph(n = 0, directed = TRUE)
  t = add_vertices(t, nv=1, attr=list(name='a'))
  t = add_vertices(t, nv=1, attr=list(name='b'))
  t = add_vertices(t, nv=1, attr=list(name='c'))
  t = add_vertices(t, nv=1, attr=list(name='d'))
  t = add_edges(t, c('a','c', 'b','c', 'c','d'))
  return(t)
}

get_nodes_names <- function(graph) {
  return(igraph::V(graph)$name)
}

add_degrees_attr <- function(graph) {
  
  nodes = get_nodes_names(graph)
  
  igraph::V(graph)$degree = igraph::degree(graph, nodes, mode='all')
  igraph::V(graph)$in_degree = igraph::degree(graph, nodes, mode='in')
  igraph::V(graph)$out_degree = igraph::degree(graph, nodes, mode='out')
  
  return(graph)
}

classify_nodes_monolri <- function(graph) {
  # TODO: Test degree attributes exist
  nodes = get_nodes_names(graph)
  types = purrr::map(
    nodes,
    function(node) {
      in_degree = igraph::V(graph)[node]$in_degree
      out_degree = igraph::V(graph)[node]$out_degree
      if (in_degree == 0) {
        return('Ligand')
      } else if (out_degree == 0) {
        return('Receptor')
      } else if (min(in_degree, out_degree) > 0) {
        return('Bifunctional')
      } else {
        stop('Node with no edges.')
      }
    }
  )
  igraph::V(graph)$type = types
  return(graph)
}

color_nodes_monolri_by_type <- function(graph) {
  # TODO: Test type attr exists
  COLOR_MAP = list(
    LIGAND_COLOR = 'red',
    BIFUNCTIONAL_COLOR = 'yellow',
    RECEPTOR_COLOR = 'blue'
  )
  
  nodes = get_nodes_names(graph)
  colors = purrr::map_chr(
    nodes, 
    function(node) {
      type = igraph::V(graph)[[node]]$type
      if (type == 'Ligand') {
        return(COLOR_MAP$LIGAND_COLOR)
      } else if (type == 'Receptor') {
        return(COLOR_MAP$RECEPTOR_COLOR)
      } else if (type == 'Bifunctional') {
        return(COLOR_MAP$BIFUNCTIONAL_COLOR)
      } else {
        stop("Node type not in Ligand, Receptor, Bifunctional")
      }
    }
  )
  igraph::V(graph)$color = colors
  return(graph)
}

is_lri_mono <- function(lri, lrdb) {
  v = lrdb[LR_GENES == lri, list(LIGAND_2, RECEPTOR_2, RECEPTOR_3)]
  return(all(is.na(v)))
}

is_node_in_graph <- function(graph, name) {
  return(name %in% igraph::V(graph)$name)
}

add_node_to_graph <- function(graph, name, type) {
  return(
    igraph::add_vertices(graph, nv=1, attr=list(name=name, type=type))
  )
}

add_edge_to_graph <- function(graph, tail, head, type) {
  return(
    igraph::add_edges(graph, c(tail, head), type=type)
  )
}

get_duplicate_nodes <- function(graph) {
  dup = V(graph)$name[duplicated(V(graph)$name)]
  return(dup)
}

# V(g)[[]]  E(g)[[]]

LIGAND_TYPE = 'LIGAND'
RECEPTOR_TYPE = 'RECEPTOR'
EDGE_TYPES = c("GENE->GENE", "GENE->COMPLEX", "COMPLEX->COMPLEX", "COMPLEX->GENE")


# LRIs = lrdb[, LR_GENES]

# Must preprocess to distinguish LRI where ligand name == receptor name
# ligands_in_identical_LRI = lrdb_mono[LIGAND_1 == RECEPTOR_1, LIGAND_1]
# receptors_in_identical_LRI = lrdb_mono[LIGAND_1 == RECEPTOR_1, RECEPTOR_1]
# lrdb_mono_ = lrdb_mono[LIGAND_1 %in% ligands_in_identical_LRI, LIGAND_1 := paste0(LIGAND_1, "__L")]
# lrdb_mono_ = lrdb_mono[LIGAND_1 %in% receptors_in_identical_LRI, RECEPTOR_1 := paste0(RECEPTOR_1, "__R")]
# LRIs = lrdb_mono_[, LR_GENES]

LRIs = lrdb_mono[, LR_GENES]
g = igraph::make_empty_graph(n = 0, directed = TRUE)
num_dup = 0
for (lri in LRIs) {
  # row = lrdb_mono_[LR_GENES == lri]
  row = lrdb_mono[LR_GENES == lri]
  if (is_lri_mono(lri, lrdb)) {
    # deal with nodes
    ligand = row[, LIGAND_1]
    receptor = row[, RECEPTOR_1]
    
    is_ligand_in_graph = is_node_in_graph(graph=g, name=ligand)
    if (!is_ligand_in_graph) {
      g = add_node_to_graph(graph=g, name=ligand, type=LIGAND_TYPE)
    }
    
    is_receptor_in_graph = is_node_in_graph(graph=g, name=receptor)
    if (!is_receptor_in_graph) {
      g = add_node_to_graph(graph=g, name=receptor, type=RECEPTOR_TYPE)
    }
    g = add_edge_to_graph(graph=g, tail=ligand, head=receptor, type="GENE->GENE")
  } else { # poly
    # deal with nodes
    # deal with edges
  }
  dup = get_duplicate_nodes(g)
  if (length(dup) > num_dup) {
    num_dup = length(dup)
    print(dup)
  }
}

# for mono, annotate nodes with L, R, Bi (bifunctional)


```
```{r}
cc_from_lrdb <- function(lrdt_mono) {}  # -> connected components

cc = igraph::components(g, mode='weak')

get_cc_subgraphs <- function(graph) {
  cc = igraph::components(graph, mode='weak')
  num_cc = cc$no
  ccs = list()
  for (i in 1:num_cc) {
    node_names = names(cc$membership)[cc$membership == i]
    ccs[[i]] = list(
      id = i,
      subgraph = igraph::induced_subgraph(graph, node_names),
      num_nodes = cc$csize[i],
      nodes = node_names
    )
  }
  return(ccs)
}

library(visNetwork)
cc_graphs = get_cc_subgraphs(g)
# visNetwork::visIgraph(cc_graphs[[5]]$subgraph)

warnings('Check ids of cc_graphs')
cc_huge = cc_graphs[[1]]$subgraph
cc_huge = add_degrees_attr(cc_huge)
cc_huge = classify_nodes_monolri(cc_huge)
cc_huge = color_nodes_monolri_by_type(cc_huge)

type_as_int = as.integer(purrr::map_dbl(
  igraph::V(cc_huge)$type, 
  function(type) {
    if (type == 'Ligand') {
      return(0)
    } else if (type == 'Bifunctional') {
      return(1)
    } else if (type == 'Receptor') {
      return(2)
    } else {
      stop('Invalid node type in LRI graph.')
    }
  }
))

# order_by_type = 
ligand_mask = igraph::V(cc_huge)$type == 'Ligand'
bifunctional_mask = igraph::V(cc_huge)$type == 'Bifunctional'
receptor_mask = igraph::V(cc_huge)$type == 'Receptor'

num_ligands = sum(ligand_mask)
num_bifuncs = sum(bifunctional_mask)
num_receptors = sum(receptor_mask)

order = 1:length(igraph::V(cc_huge))
order[ligand_mask] = 1:num_ligands
order[bifunctional_mask] = (1:num_bifuncs) + num_ligands
order[receptor_mask] = (1:num_receptors) + num_ligands + num_bifuncs

v_order = igraph::V(cc_huge)[order]

layout_hierarchical = igraph::layout_with_sugiyama(
  cc_huge, 
  layers = type_as_int, 
  hgap = 1, vgap = 1,
  maxiter = 5)

layout_circle = igraph::layout_in_circle(cc_huge, order=v_order)

```

Huge cc with 1615 nodes (id = 1)
Tac1/2/4 - neurohormones (cc id=32) - ok
Dsg3 - cell junction / ECM interactions (id=8)
Guanylate cyclase activator and guanylate cyclase (id=42) - ?
IL22 (id=45) - ok
Lipase H (id=50) - ok


# Celltype LR graphs

```{r}

cci_det = cci_detected

res = purrr::pmap(
  cci_det,
  function(
    EMITTER_CELLTYPE, RECEIVER_CELLTYPE,
    LIGAND_1, LIGAND_2,
    RECEPTOR_1, RECEPTOR_2, RECEPTOR_3,
    IS_CCI_DETECTED_YOUNG, IS_CCI_DETECTED_OLD,
    ...) {
    
    
    ligands_young = c()
    ligands_old = c()
    receptors_young = c()
    receptors_old = c()
    
    if (IS_CCI_DETECTED_YOUNG) {
      ligands_young = c(ligands_young, LIGAND_1)
      ligands_young = c(ligands_young, LIGAND_2)
      receptors_young = c(receptors_young, RECEPTOR_1)
      receptors_young = c(receptors_young, RECEPTOR_2)
      receptors_young = c(receptors_young, RECEPTOR_3)
    }
    if (IS_CCI_DETECTED_OLD) {
      ligands_old = c(ligands_old, LIGAND_1)
      ligands_old = c(ligands_old, LIGAND_2)
      receptors_old = c(receptors_old, RECEPTOR_1)
      receptors_old = c(receptors_old, RECEPTOR_2)
      receptors_old = c(receptors_old, RECEPTOR_3)
    }
    
    res = list(
      EMITTER_CELLTYPE = EMITTER_CELLTYPE,
      RECEIVER_CELLTYPE = RECEIVER_CELLTYPE,
      LIGANDS_DETECTED_YOUNG = ligands_young,
      RECEPTORS_DETECTED_YOUNG = receptors_young,
      LIGANDS_DETECTED_OLD = ligands_old,
      RECEPTORS_DETECTED_OLD = receptors_old
    )
        
    return(res)
    
    }
  )


# to dt
library(data.table)
res_ = as.data.frame(do.call(rbind, res))
setDT(res_)

# unlist (for grouping in data.table)
for (col in c("EMITTER_CELLTYPE", "RECEIVER_CELLTYPE")) {
  res_[, eval(col) := unlist(get(col), recursive=FALSE)]
}

from_emitters = res_[, list(
    CELLTYPE = EMITTER_CELLTYPE,
    LIGANDS_DETECTED_YOUNG=list(LIGANDS_DETECTED_YOUNG),
    LIGANDS_DETECTED_OLD=list(LIGANDS_DETECTED_OLD)
  ), 
  by=EMITTER_CELLTYPE]

from_receivers = res_[, list(
    CELLTYPE = RECEIVER_CELLTYPE,
    RECEPTORS_DETECTED_YOUNG=list(RECEPTORS_DETECTED_YOUNG),
    RECEPTORS_DETECTED_OLD=list(RECEPTORS_DETECTED_OLD)
  ), 
  by=RECEIVER_CELLTYPE]

celltypes_comm = merge(from_emitters, from_receivers,
                  by='CELLTYPE',
                  all=FALSE)

celltypes_comm[, EMITTER_CELLTYPE := NULL]
celltypes_comm[, RECEIVER_CELLTYPE := NULL]

list2chr <- function(l) {unique(unlist(l))}
celltypes_comm[, LIGANDS_DETECTED_YOUNG := map(celltypes_comm[, LIGANDS_DETECTED_YOUNG], list2chr)]
celltypes_comm[, LIGANDS_DETECTED_OLD := map(celltypes_comm[, LIGANDS_DETECTED_OLD], list2chr)]
celltypes_comm[, RECEPTORS_DETECTED_YOUNG := map(celltypes_comm[, RECEPTORS_DETECTED_YOUNG], list2chr)]
celltypes_comm[, RECEPTORS_DETECTED_OLD := map(celltypes_comm[, RECEPTORS_DETECTED_OLD], list2chr)]

celltypes_comm[, NUM_LIGANDS_YOUNG := map_int(celltypes_comm[, LIGANDS_DETECTED_YOUNG], length)]
celltypes_comm[, NUM_LIGANDS_OLD := map_int(celltypes_comm[, LIGANDS_DETECTED_OLD], length)]
celltypes_comm[, NUM_RECEPTORS_YOUNG := map_int(celltypes_comm[, RECEPTORS_DETECTED_YOUNG], length)]
celltypes_comm[, NUM_RECEPTORS_OLD := map_int(celltypes_comm[, RECEPTORS_DETECTED_OLD], length)]

celltypes_comm
```

Plot celltype communication change using parallel coordinates chart
```{r}
library(ggplot2)
library(GGally)
library(ggpubr)

cell_counts <- function(cci_detected) {
  counts = cci_det[
    , 
    list(
      NCELLS_YOUNG = unique(NCELLS_EMITTER_YOUNG),
      NCELLS_OLD = unique(NCELLS_EMITTER_OLD)
      ), 
    by = EMITTER_CELLTYPE
  ]
  
  counts[, CELLTYPE := EMITTER_CELLTYPE]
  counts[, EMITTER_CELLTYPE := NULL]
  return(counts)
}

build_parcoord_plot <- function(dt, col_idxs, col_group, title) {
  return(
    ggparcoord(
      dt,
      columns = col_idxs, 
      groupColumn = col_group,
      scale = 'globalminmax',
      # scaleSummary = 'mean',
      showPoints = TRUE,
      title = title,
      alphaLines = 0.7
      ) + theme(plot.title = element_text(size=12)) +
      xlab('') + ylab('Counts')
  )
}

g1 = ggparcoord(celltypes_comm,
    columns = 6:7, 
    groupColumn = 1,
    scale = 'globalminmax',
    # scaleSummary = 'mean',
    showPoints = TRUE,
    title = 'Detected genes of ligands participating in CCIs',
    alphaLines = 0.7
    ) + theme(plot.title = element_text(size=12)) +
    xlab('') + ylab('Counts')
g2 = ggparcoord(celltypes_comm,
    columns = 8:9, 
    groupColumn = 1,
    scale = 'globalminmax',
    # scaleSummary = 'mean',
    showPoints = TRUE,
    title = 'Detected genes of receptors participating in CCIs',
    alphaLines = 0.7
    ) + theme(plot.title = element_text(size=12)) +
    xlab('') + ylab('Counts')
g3 = ggparcoord(cell_counts(cci_det),
    columns = 1:2, 
    groupColumn = 3,
    scale = 'globalminmax',
    # scaleSummary = 'mean',
    showPoints = TRUE,
    title = 'Cells sequenced',
    alphaLines = 0.7
    ) + theme(plot.title = element_text(size=12)) +
    xlab('') + ylab('Counts')

ggarrange(g1, g2, g3,
          nrow=3, ncol=1)
```

```{r}
library(ComplexUpset)

col = 'RECEPTORS_DETECTED_OLD'
genes = unique(
  purrr::reduce(celltypes_comm[, get(col)], union)
)
celltypes = celltypes_comm[, CELLTYPE]

bool = map(genes, 
    function(x, celltypes, celltypes_comm, col) {
      res = list()
      for (cell in celltypes) {
        genes = celltypes_comm[CELLTYPE == cell, get(col)][[1]]
        res[cell] = (x %in% genes)
      }
      return(as.logical(res))
    }, celltypes, celltypes_comm, col)
names(bool) = genes

setDT(bool)
# bool[, CELLTYPE := celltypes]

bool_t = data.table::transpose(bool)
names(bool_t) = celltypes
bool_t_df = as.data.frame(bool_t)

upset(data=bool_t_df, intersect=celltypes, min_size=5, wrap=FALSE) + ggtitle(col)
# ggarrange(x, x, nrow=2, ncol=1)
```

# Celltype pairs graphs

```{r}
LRI_by_pair_young = cci_det[IS_CCI_DETECTED_YOUNG == TRUE, list(LRI_YOUNG = list(LR_GENES)), by=list(EMITTER_CELLTYPE, RECEIVER_CELLTYPE)]
LRI_by_pair_old = cci_det[IS_CCI_DETECTED_OLD == TRUE, list(LRI_OLD = list(LR_GENES)), by=list(EMITTER_CELLTYPE, RECEIVER_CELLTYPE)]

LRI_by_pair = merge(LRI_by_pair_young, LRI_by_pair_old, by=c('EMITTER_CELLTYPE', 'RECEIVER_CELLTYPE'), all=FALSE)

LRI_by_pair[, Num_LRI_YOUNG := map_int(LRI_by_pair[, LRI_YOUNG], length)]
LRI_by_pair[, Num_LRI_OLD := map_int(LRI_by_pair[, LRI_OLD], length)]
LRI_by_pair[, ER_CELLTYPE := paste0(EMITTER_CELLTYPE, "__", RECEIVER_CELLTYPE)]
LRI_by_pair
```

```{r}
build_parcoord_plot(LRI_by_pair, 5:6, 7, 'Number of LRIs by cell pairs')
```

```{r}
col = 'LRI_YOUNG'

LRI = unique(
  purrr::reduce(LRI_by_pair[, get(col)], union)
)

celltype_pairs = LRI_by_pair[, ER_CELLTYPE]
bool = map(LRI,
    function(x, celltype_pairs, LRI_by_pair, col) {
      res = list()
      for (cell_pair in celltype_pairs) {
        lri = LRI_by_pair[ER_CELLTYPE == cell_pair, get(col)][[1]]
        res[cell_pair] = (x %in% lri)
      }
      return(as.logical(res))
    }, celltype_pairs, LRI_by_pair, col)
names(bool) = LRI

setDT(bool)

bool_t = data.table::transpose(bool)
names(bool_t) = celltype_pairs
bool_t_df = as.data.frame(bool_t)

upset(data=bool_t_df, intersect=celltype_pairs, min_size=5, wrap=FALSE) + ggtitle(col)
# # ggarrange(x, x, nrow=2, ncol=1)

# celltype_pairs_sel = LRI_by_pair[RECEIVER_CELLTYPE == 'endothelial cell of hepatic sinusoid', ER_CELLTYPE]
celltype_pairs_sel = LRI_by_pair[EMITTER_CELLTYPE == 'endothelial cell of hepatic sinusoid', ER_CELLTYPE]
upset(data=bool_t_df, intersect=celltype_pairs_sel, min_size=5, wrap=FALSE) + ggtitle(col)

```

General CCI change graph

```{r}

get_complete_cci_change_graph <- function(cci_detected) {
  dt_edge = cci_detected[REGULATION %in% c('UP', 'DOWN'), 
        list(
          EMITTER_CELLTYPE, RECEIVER_CELLTYPE, 
          LR_GENES, LOGFC,
          REGULATION
          )
        ][,
          value := abs(LOGFC)][
            ,
            label := format(round(LOGFC, 2), nsmall=2)][
              ,
              color := ifelse(REGULATION == 'UP', 'red', 'blue')
            ]
  cci_change_graph = igraph::graph_from_data_frame(dt_edge, directed = TRUE, vertices = NULL)
  return(cci_change_graph)
}
LRI_subgraph <- function(cci_detected, LRIs) {
  SUBG_LIMIT = 5
  if(length(LRIs) > SUBG_LIMIT) {
    stop('The number of subgraphs limit for visualization is SUBG_LIMIT')
  }
  dt_edge = cci_detected[
    (REGULATION %in% c('UP', 'DOWN'))
      & (LR_GENES %in% LRIs),
        list(
          EMITTER_CELLTYPE, RECEIVER_CELLTYPE, 
          LR_GENES, LOGFC,
          REGULATION
          )
        ][,
          value := abs(LOGFC)][
            ,
            # label := format(round(LOGFC, 2), nsmall=2)][
            label := LR_GENES][
              ,
              color := ifelse(REGULATION == 'UP', 'red', 'blue')
            ]
  g = igraph::graph_from_data_frame(dt_edge, directed = TRUE, vertices = NULL)
  g$LRI = LRIs
  return(g)
}
build_interactive_LRI_network <- function(cci_detected, LRIs) {
  g = LRI_subgraph(cci_detected, LRIs)
  vis = visIgraph(g) %>%
    visEdges(font='8px arial green',
             smooth=list(roundness=1),
             scaling=list(
               min=10, max=16,
               label=list(min=14, max=20))) %>%
    ( scDiffCom:::get_visnetwork_options() ) %>%
    ( scDiffCom:::get_visnetwork_interactive() )
  return(vis)
}

get_LRI_subgraphs <- function(cci_detected, LRIs = NULL) {
  if(is.null(LRIs)) {
    lris = unique(cci_detected[, LR_GENES])
  } else {
    lris = LRIs
  }
  subgraphs = purrr::map(
    lris,
    ~ LRI_subgraph(cci_detected, .x)
  )
  if(length(lris)==1) {
    return(subgraphs[[1]])
  } else {
    return(subgraphs)
  }
}
LRI_subgraph_metrics <- function(cci_detected, LRIs=NULL) {
  subgraphs = get_LRI_subgraphs(cci_detected, LRIs)
  
  subg_metrics_l = purrr::map(
    subgraphs,
    function(subg) {
      
      metrics = list(
          LRI = subg$LRI,
          num_edges = length(igraph::E(subg)),
          num_loops = sum(igraph::which_loop(subg)),
          num_reciprocal = sum(igraph::which_mutual(subg)) - sum(igraph::which_loop(subg)),
          motifs_3_total = igraph::count_motifs(subg, size=3),
        )
      motifs_3 = as.list(igraph::motifs(subg, size = 3))
      names(motifs_3) = paste0('Motifs_3_id_', 0:15)
      
      return(c(metrics, motifs_3))
  
    }
  )

  subg_metrics_dt = data.table::rbindlist(subg_metrics_l)
  return(subg_metrics_dt)
}

get_interesting_LRI_graphs <- function(subg_metrics, 
                                          cci_detected=NULL,
                                          isom_class=NULL,
                                          byNumEdges=FALSE) {
  if(!xor(is.null(subg_metrics), is.null(cci_detected))) {
    stop('subg_metrics XOR cci_detected must be provided')
  }
  if(!xor(is.null(isom_class), byNumEdges)) {
    stop('isom_class XOR byNumEdges must be specified')
  }
  if(!is.null(cci_detected)) {
    subg_metrics = LRI_subgraph_metrics(cci_detected)
  }
  
  if(!is.null(isom_class)) {
    colname = paste0('Motifs_3_id_', isom_class)
    lris = subg_metrics[get(colname) > 0][order(-get(colname)), LRI]
    return(lris)
  } else {
    lris_sorted = subg_metrics[order(-num_edges), LRI]
    return(lris_sorted)
  }
  
}

# Get general CCI change graph
cci_change_graph = get_complete_cci_change_graph(cci_det)
# igraph::V(cci_change_graph)$smooth = TRUE

# LRI = "Ubb:Ripk1"
LRI = "Gcg:Adrb2"

cci_change_subg = get_LRI_subgraphs(cci_det, LRI)
cci_change_subg
visIgraph(cci_change_subg)
# visIgraph(cci_change_graph)


# Get every subgraph and compute metrics of interest
LRIs = unique(igraph::E(cci_change_graph)$LR_GENES)

subg_metrics_dt = LRI_subgraph_metrics(cci_det)

purrr::map(
  names(subg_metrics_dt)[2:ncol(subg_metrics_dt)],
  function(col) {
    print(col)
    if (!is.na(sum(subg_metrics_dt[, get(col)]))) {
      return( hist(subg_metrics_dt[[col]], xlab=col) )
    }
    else {
      return(NULL)
    }
  }
)

```
```{r}
# Visualize 3-vertex graph isomorphism classes
map(0:15, ~ plot(graph_from_isomorphism_class(3, .x), main=paste0('isomorphism class = ', .x)))
```

```{r}
subg_metrics_dt
```

Technical: Association between min(ncells) and |logFC| and p-value

```{r}
min_cells_young = purrr::map2_dbl(cci_det[, NCELLS_EMITTER_YOUNG], cci_det[, NCELLS_RECEIVER_YOUNG], min)
min_cells_old = purrr::map2_dbl(cci_det[, NCELLS_EMITTER_OLD], cci_det[, NCELLS_RECEIVER_OLD], min)
min_cells = purrr::map2_dbl(min_cells_young, min_cells_old, min)
logFC = cci_det[["LOGFC"]]
abslogFC = cci_det[["LOGFC_ABS"]]
p_val = cci_det[["P_VALUE_DE"]]

x = min_cells
y = p_val
{
  plot(x, y)
  abline(lm(y ~ x), col='red')
}
```

