---
title: "R Notebook"
---

Cell-ontology analysis and cell-type annotation of TMS and Calico datasets.

First we load the csv file that contains all the cell types and the annotations.

```{r}
library(data.table)
DATASET_PATH <- "../../data_scAgeCom/analysis/scDiffCom_cell_types.csv"

cell_types_dt <- setDT(read.csv(DATASET_PATH, stringsAsFactors = FALSE))
  
message("Dataset composition:")
cell_types_dt

message("Datase columns:")
colnames(cell_types_dt)

```

# Cell-types and families

Look at the content of the table. Note that for now some annotations are unsure, so we just need to clean a bit (e.g. remove special characters)

```{r}
message("Extract all tissues:")
tissues <- unique(cell_types_dt$Tissue)
head(tissues)

message("Extract cell-types as names in the original studies:")
cell_types_original <- unique(cell_types_dt$TMS.Calico.cell.type)
head(cell_types_original)
tissue_cell_types_original <- unique(paste(cell_types_dt$Tissue, cell_types_dt$TMS.Calico.cell.type, sep = "_"))
head(tissue_cell_types_original)

message("Extract cell-types as renamed for scDiffCom analysis:")
cell_types_scdiffcom <- unique(cell_types_dt$scDiffCom.cell.type)
head(cell_types_scdiffcom)
tissue_cell_types_scdiffcom <- unique(paste(cell_types_dt$Tissue, cell_types_dt$scDiffCom.cell.type, sep = "_"))
head(tissue_cell_types_scdiffcom)

message("Extract familiy annotation")
family_annotation <- unique(cell_types_dt$Family...broad)
family_annotation <- family_annotation[!(family_annotation == "" | grepl("\\+|\\?| or |\\,", family_annotation))]
sort(family_annotation)

```

# Cell ontology generalities

We'll look at the cell ontology.

General purpose ontology manipulation in R can be done with ontologyX tools (ontologyIndex, ontologyPlot, ontologySimilarity?) and ontoProc offers bio-geared ontology functionality.

```{r}
library(ontoProc)
library(ontologyIndex)
library(ontologyPlot)
library(purrr)
```

General ontologies are often stored in .owl format and, in terms of technologies, ontologies can be served by graph databases. OBO, the open biological ontologies foundation, offers bio-related ontologies in .OBO format:
http://www.obofoundry.org/ontology/cl.html

getCellOnto() can be used to extract ontology and represent it as orthology_index class as defined by the orthologyIndex package. orthologyIndex is a package for general-purpose low-level ontology manipulation.

Let's explore oncoProc functionality
```{r}
cl = getCellOnto()  # just be careful with the version of the package/database
# cl = get_ontology(file='../../data/cl.obo')
cl_names <- cl$name

cl$name['CL:0000127']

# Plot CL for terms
cell_v = c("CL:0000492", "CL:0001054", "CL:0000236", "CL:0000625",
   "CL:0000576", "CL:0000623", "CL:0000451", "CL:0000556")
onto_plot2(cl, cell_v)

# feats = CLfeats(cl, "CL:0000492")

# Names from cell line ontology
# cleannames = cleanCLOnames()

# Start app and dataframes with cell types properties
# ctmarks(cl)


# onto_roots(cl)
recognizedPredicates()
```

Explore ontologyIndex functionality
```{r}
# Explore the data structure
str(cl, max.level=1)
term = "CL:0000492"
cl$name[term]
cl$parents[term]
cl$name[cl$ancestors[[term]]]

# Get term properties
get_term_property(
    ontology=cl, 
    property="ancestors",
    term="CL:0000492",
    as_names=TRUE
)

# Remove redundant terms from a set of terms
terms = c("CL:0000492", "CL:0001054", "CL:0000236", "CL:0000625",
   "CL:0000576", "CL:0000623")
minimal = minimal_set(cl, terms)

# Find all ancestors of a set of terms
ancestors_ids = get_ancestors(cl, terms)
cl$name[ancestors_ids]

# Operations on subclasses
# intersection_with_descendants()
# exclude_descendants()
# prune_descendants()

properties = names(cl)
# cl$has_high_plasma_membrane_amount

```


# Below: to refresh and add parts from other scripts



```{r}

# -------- function to plot cell types per tissue and family
plot_celltypes <- function(dt, tissue=tissues, family=family_annotation, action="save"){
  our_tissue_cell_types <- subset(dt, Tissue %in% tissue & Family...broad %in% family)
  #our_tissue_cell_types <- unique(c(our_tissue_cell_types$scDiffCom.cell.type,
  #                                  our_tissue_cell_types$TMS.Calico.cell.type))
  # to use only scDiffCom cell types, comment line above and uncomment line below
  our_tissue_cell_types <- unique(our_tissue_cell_types$scDiffCom.cell.type)
  
  tissue_names <- cl_names[cl_names %in% our_tissue_cell_types]
  
  if(action == "show"){
    onto_plot2(cl, names(tissue_names), cex = 0.8)
  } else {
    if(length(tissue) == 1 & length(family) == 1){
      png(file=paste0("Plots/Mixed/",tissue,"_",family,".png"),width=2000, height=1000)
      onto_plot2(cl, names(tissue_names), cex = 0.8)
      dev.off()
    } else if(length(family) == 1 & identical(tissue,all_tissues)) {
      png(file=paste0("Plots/Families/",family,".png"),width=2000, height=1000)
      onto_plot2(cl, names(tissue_names), cex = 0.8)
      dev.off()
    } else if(length(tissue) == 1 & identical(family,all_cellfamily)) {
      png(file=paste0("Plots/Tissues/",tissue,".png"),width=2000, height=1000)
      onto_plot2(cl, names(tissue_names), cex = 0.8)
      dev.off()
    } else {
      onto_plot2(cl, names(tissue_names), cex = 0.8)
      message("Please save plot manually to give it a proper name")
    }
  }
}


# -------- Using function

# Plot and show all cell types for one tissue
plot_celltypes(tissue = "Liver", action = "show")

# Plot and save one cell type for one tissue
plot_celltypes(tissue = "Spleen", family = "leukocyte")

# Plot and save all cell types for each tissue
sapply(all_tissues, function(x){
  plot_celltypes(tissue = x)
}
)
# Plot and save each cell type families for all tissues
sapply(all_cellfamily, function(x){
  plot_celltypes(family = x)
}
)


```


