---
title: "R Notebook"
---

```{r}
library(data.table)
library(purrr)
library(scDiffCom)
library(ontologyIndex)
library(reshape2)
library(ggplot2)
library(ggpubr)
library(ontologyPlot)
# library(GO.db)

obj = readRDS('../../data_scAgeCom/testing/scDiffCom_test_facs_liver.rds')
GO = ontologyIndex::get_ontology(file='../../data/go.obo')
LR_GO = scDiffCom::LRdb_mouse$LRdb_curated_GO

BIOLOGICAL_PROCESS_ROOT = 'GO:0008150'
MOLECULAR_FUNCTION_ROOT = 'GO:0003674'
CELLULAR_COMPONENT_ROOT = 'GO:0005575'

BIOLOGICAL_PROCESSES = ontologyIndex::get_descendants(GO, BIOLOGICAL_PROCESS_ROOT)
MOLECULAR_FUNCTIONS = ontologyIndex::get_descendants(GO, MOLECULAR_FUNCTION_ROOT)
CELLULAR_COMPONENTS = ontologyIndex::get_descendants(GO, CELLULAR_COMPONENT_ROOT)

get_GO <- function() {
  return(
    ontologyIndex::get_ontology(file='../../data/go.obo')
  )
}
```

Utils

```{r}
# (A)
build_GO_LRI_binary_indicator_dt <- function() {
  lr_go = scDiffCom::LRdb_mouse$LRdb_curated_GO
  lr_go[, 'Indicator' := 1]
  lr_go[, 'GO_NAME' := NULL]
  lr_go_wide = data.table::dcast(lr_go, LR_SORTED ~ GO_ID, value.var = 'Indicator')
  lr_go_wide[is.na(lr_go_wide)] = 0
  rownames(lr_go_wide) = lr_go_wide[["LR_SORTED"]]
  lr_go_wide[, LR_SORTED := NULL]
  return(lr_go_wide)  
  # head(lr_go_wide)
  # sum(lr_go_wide[, 2:3641])
}
plot_GO_LRI_histogram <- function(lr_go_wide = NULL, by_GO_category = FALSE) {
  
  if(is.null(lr_go_wide)) {
    lr_go_wide = build_GO_LRI_binary_indicator_dt()
  }
  
  hist_from_vector <- function(v, bw=NULL, xlim=NULL, xlab=NULL, title=NULL) {
      return(
        ggplot(data.frame(v), aes(v))
        + geom_histogram(binwidth=bw)
        + coord_cartesian(xlim=xlim)
        + xlab(xlab)
        + ggtitle(label=title)
      )
    }
  
  Num_LRIs = dim(lr_go_wide)[1]
  Num_GOs = dim(lr_go_wide)[2]
  
  if(!by_GO_category) {

    Num_GOs_per_LR = rowSums(lr_go_wide)
    Num_LRIs_per_GO = colSums(lr_go_wide)
    
    g1 = hist_from_vector(Num_GOs_per_LR,
                          bw=5,
                          xlab='#GO per LR',
                          title=paste('#GOs = ', Num_GOs))
    g2 = hist_from_vector(Num_LRIs_per_GO,
                          bw=10, 
                          xlim=c(0, 500),
                          xlab='#LRI per GO term',
                          title=paste('#LRIs = ', Num_LRIs))
    gg = ggarrange(g1, g2,
              nrow=1, ncol=2)
    return(gg)
  
  } else {
    
    BIOLOGICAL_PROCESSES = ontologyIndex::get_descendants(GO, BIOLOGICAL_PROCESS_ROOT)
    MOLECULAR_FUNCTIONS = ontologyIndex::get_descendants(GO, MOLECULAR_FUNCTION_ROOT)
    CELLULAR_COMPONENTS = ontologyIndex::get_descendants(GO, CELLULAR_COMPONENT_ROOT)
    
    BPs = colnames(lr_go_wide)[colnames(lr_go_wide) %in% BIOLOGICAL_PROCESSES]
    MFs = colnames(lr_go_wide)[colnames(lr_go_wide) %in% MOLECULAR_FUNCTIONS]
    CCs = colnames(lr_go_wide)[colnames(lr_go_wide) %in% CELLULAR_COMPONENTS]
    
    Num_BPs = length(BPs)
    Num_MFs = length(MFs)
    Num_CCs = length(CCs)
    
    Num_BPs_per_LR = rowSums(lr_go_wide[, ..BPs])
    Num_MFs_per_LR = rowSums(lr_go_wide[, ..MFs])
    Num_CCs_per_LR = rowSums(lr_go_wide[, ..CCs])
    Num_LRIs_per_BPs = colSums(lr_go_wide[, ..BPs])
    Num_LRIs_per_MFs = colSums(lr_go_wide[, ..MFs])
    Num_LRIs_per_CCs = colSums(lr_go_wide[, ..CCs])
    
    g1 = hist_from_vector(Num_BPs_per_LR,
                          bw=5,
                          xlab='#BPs per LRI',
                          title=paste('#BPs = ', Num_BPs))
    g2 = hist_from_vector(Num_LRIs_per_BPs,
                          bw=10, 
                          xlim=c(0, 500),
                          xlab='#LRI per BP',
                          title=paste('#LRIs = ', Num_LRIs))
    g3 = hist_from_vector(Num_MFs_per_LR,
                          bw=5,
                          xlab='#MFs per LRI',
                          title=paste('#MFs = ', Num_MFs))
    g4 = hist_from_vector(Num_LRIs_per_MFs,
                          bw=10, 
                          xlim=c(0, 500),
                          xlab='#LRI per MF',
                          title=paste('#LRIs = ', Num_LRIs))
    g5 = hist_from_vector(Num_CCs_per_LR,
                          bw=5,
                          xlab='#CCs per LRI',
                          title=paste('#CCs = ', Num_CCs))
    g6 = hist_from_vector(Num_LRIs_per_CCs,
                          bw=10, 
                          xlim=c(0, 500),
                          xlab='#LRI per CC',
                          title=paste('#LRIs = ', Num_LRIs))
    gg = ggarrange(
      g1, g2, g3, g4, g5, g6,
      nrow=3, ncol=2)
    return(gg)
    
  }
  
}

# (B)
# Will approximate
get_term_num_ancestors <- function(go, terms) {
    return(
        map_int(
            terms,
            ~ (length(get_ancestors(go, .x)))
        ) - 1
    )
}
summarize_GO_terms_landscape <- function(ora_dt, go_id_colname, gene_ontology) {
  
  ora_dt_ext = extend_GO_terms_dt(ora_dt, go_id_colname, gene_ontology)
  
  freqs = ora_dt_ext[['FREQUENCY']]
  num_ancestors = ora_dt_ext[['NUM_ANCESTORS']]

  plot_hist <- function(vec, xlab, bw=0.01, zoom=FALSE) {
      if(!zoom) {
          return(
              ggplot(data.frame(vec), aes(vec))
              + geom_histogram(binwidth=bw)
              + xlab(xlab)
              + theme_bw()
          )
      } else {
          return(
              ggplot(data.frame(vec), aes(vec))
              + geom_histogram(binwidth=bw)
              + xlab(xlab)
              + coord_cartesian(xlim=c(0, 0.05), ylim=c(0, 100))
              + theme_bw()
          )
      }
  }
  freq_hist = plot_hist(freqs, 'Term frequency', 0.001, zoom=TRUE)
  ancest_hist = plot_hist(num_ancestors, 'Num ancestors', 1)

  plot_data = data.table(
    GO=names(freqs),
    TERM_FREQ=freqs,
    NUM_ANCESTORS=num_ancestors
  )
  
  plot_scatter <- function(data, zoom=FALSE) {
    if (!zoom) {
        return(
            ggplot(data, aes(x=NUM_ANCESTORS, y=TERM_FREQ))
            + geom_jitter()
            + geom_hline(yintercept=0.01)
            + geom_hline(yintercept=0.025)
            + geom_hline(yintercept=0.05)
            + theme_bw()
        )
    }
    else {
        return(
            ggplot(data, aes(x=NUM_ANCESTORS, y=TERM_FREQ))
            + geom_jitter()
            + coord_cartesian(xlim=c(0, 20), ylim=c(0, 0.10))
            + geom_hline(yintercept=0.01)
            + geom_hline(yintercept=0.025)
            + geom_hline(yintercept=0.05)
            + theme_bw()
        )
    }
    
}
  unzoom_scatter = plot_scatter(plot_data, zoom=FALSE)
  zoomed_scatter = plot_scatter(plot_data, zoom=TRUE)

  g = ggarrange(
    freq_hist, ancest_hist,
    unzoom_scatter, zoomed_scatter,
    nrow=2, ncol=2)
  
  return(
    annotate_figure(g)
  )
}
describe_GO_terms_set <- function(gos, gene_ontology) {
    
  freqs = get_term_frequencies(gene_ontology, gos)
  freqs = freqs[ names(freqs) %in% gos ]  # for some reason increases
  num_ancestors = get_term_num_ancestors(gene_ontology, gos)
  info_content = get_term_info_content(gene_ontology, gos)
  info_content = info_content[ names(info_content) %in% gos ]  # for some reason increases
    
  # names(freqs) = gos
  names(num_ancestors) = gos
  # names(info_content) = gos
  
  # NA-fills up to length(gos)
  length(freqs) <- length(info_content) <- length(num_ancestors) <- length(gos)
  
  dt = data.table(
    GO_TERM = gos,
    FREQUENCY = freqs,
    NUM_ANCESTORS = num_ancestors,
    INFO_CONTENT = info_content
  )
  
  return(dt)
}
extend_GO_terms_dt <- function(dt_go, go_id_colname='GO_TERM', gene_ontology=GO) {
  
  gos = dt_go[[go_id_colname]]
  if (length(unique(gos)) != length(gos)) {
    stop("No duplicated GO terms are allowed")
  }
  
  dt_go_extension = describe_GO_terms_set(gos, gene_ontology)
  dt_extended = merge(dt_go, dt_go_extension,
                      by.x = go_id_colname,
                      by.y = "GO_TERM",
                      all = FALSE)
  
  return(dt_extended)
}

# (C)
plot_ORA_histograms <- function(ora_dt) {
  
  or_diff = ora_dt[, .(OR_DIFF)]
  pval_diff = ora_dt[, .(BH_P_VALUE_DIFF)]
        
  or_up = ora_dt[, .(OR_UP)]
  pval_up = ora_dt[, .(BH_P_VALUE_UP)]
        
  or_down = ora_dt[, .(OR_DOWN)]
  pval_down = ora_dt[, .(BH_P_VALUE_DOWN)]
  
  g_OR_DIFF = (
    ggplot(or_diff, aes(OR_DIFF))
    + geom_histogram(binwidth=1)
    + geom_vline(aes(xintercept=10, colour='red'))
    + ggtitle('OR_DIFF')
  )
  g_pval_DIFF = (
    ggplot(pval_diff, aes(BH_P_VALUE_DIFF))
    + geom_histogram(binwidth=0.01)
    + geom_vline(aes(xintercept=0.05, colour='red'))
    + ggtitle('BH_P_VALUE_DIFF')
  )
        
  g_OR_UP = (
    ggplot(or_up, aes(OR_UP))
    + geom_histogram(binwidth=1)
    + geom_vline(aes(xintercept=10, colour='red'))
    + ggtitle('OR_UP')
  )
  g_pval_UP = (
    ggplot(pval_up, aes(BH_P_VALUE_UP))
    + geom_histogram(binwidth=0.01)
    + geom_vline(aes(xintercept=0.05, colour='red'))
    + ggtitle('BH_P_VALUE_UP')
  )
        
  g_OR_DOWN = (
    ggplot(or_down, aes(OR_DOWN))
    + geom_histogram(binwidth=1)
    + geom_vline(aes(xintercept=10, colour='red'))
    + ggtitle('OR_DOWN')
  )
  g_pval_DOWN = (
    ggplot(pval_down, aes(BH_P_VALUE_DOWN))
    + geom_histogram(binwidth=0.01)
    + geom_vline(aes(xintercept=0.05, colour='red'))
    + ggtitle('BH_P_VALUE_DOWN')
  )
        
  ggarrange(g_OR_DIFF, g_pval_DIFF,
            g_OR_UP, g_pval_UP,
            g_OR_DOWN, g_pval_DOWN,
            nrow=3, ncol=2)
}

```



(A) Histogram of #GO per LRI and #LRI per GO

```{r}
plot_GO_LRI_histogram(by_GO_category = FALSE)
plot_GO_LRI_histogram(by_GO_category = TRUE)
```

(B) Extract num_ancestors and frequency of a GO set

> num_ancestors <= 0
GO:0008150 GO:0044420 GO:0044422 GO:0044431 GO:0044446 GO:0044456 GO:0044464 GO:0097458 GO:0098589 
         0          0          0          0          0          0          0          0         -1 

> sort(num_ancestors)[(length(num_ancestors)-10):length(num_ancestors)]
GO:0050731 GO:0010524 GO:0045773 GO:0006469 GO:0045860 GO:0043409 GO:0071901 GO:0071902 GO:0051281 GO:0043406 GO:0043407
        30         35         36         38         38         40         40         40         46         54         54 

> sort(num_ancestors)[1:20]
GO:0098589 GO:0008150 GO:0044420 GO:0044422 GO:0044431 GO:0044446 GO:0044456 GO:0044464 GO:0097458 GO:0003824 GO:0005215 GO:0008152 GO:0009987 
        -1          0          0          0          0          0          0          0          0          1          1          1          1 
GO:0022610 GO:0032501 GO:0040007 GO:0040011 GO:0050896 GO:0051179 GO:0001775 
         1          1          1          1          1          1          2 
         
```{r}
ora_dt = obj@ora_default$GO_TERMS
ora_dt_select = ora_dt[(OR_UP > 3) & (BH_P_VALUE_UP < 0.05)]

ora_dt_select_extended = extend_GO_terms_dt(ora_dt_select, 'VALUE_BIS', GO)
summarize_GO_terms_landscape(ora_dt_select, 'VALUE_BIS', GO)

head(ora_dt_select_extended)
```

(C) OR_{UP, DOWN, DIFF}|pval_{UP, DOWN, DIFF} distributions for an ORA set (not necessarily GO terms, any)

```{r}
plot_ORA_histograms(obj@ora_default$GO_TERMS)
plot_ORA_histograms(obj@ora_default$LR_GENES)
plot_ORA_histograms(obj@ora_default$ER_CELLTYPES)
```

(D) Correlations among GO terms

```{r}

```